Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-06-30 summary.Rmd

@@

>>>>

```{r lambda_est_p, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue). Simulations were repeated 50 times for each sample size. Communities were drawn to be approximately balanced.', fig.width = 10}
rmse.df <- readr::read_csv('rmse-p.csv')

@@

```{r lambda_est_p_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue). Simulations were repeated 50 times for each sample size. Communities were drawn to be imbalanced.', fig.width = 10}
rmse.df <- readr::read_csv('rmse-p-imbalanced.csv')

### 2021-06-30 summary.tex

Define the subspace incoherence:

$$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$

Then $\forall k, n$,

@@

\emph{Proof}. Since by Theorem \ref{theorem2} each \(\mathcal{S}^{(k)}\)
are mutually orthogonal, any vector projected to \(\mathcal{S}^{(k)}\)
must be orthogonal to each row of \(V^{(-k)}\). Therefore,
\(W^{(k)} v = 0\) \(\forall v \in \mathcal{S}^{(-k)}\).

@@

\begin{remark}
Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $||v_n^{(i)}|| = 1$,
which scales each $r_n^{(k)} \leq 1$. This is not strictly necessary for
the proof of Theorem \ref{theorem5} since each $\mu_n^{(k)} = 0$, so as long
as the $k^{th}$ community spans its subspace, $a r_n^{(k)} > 0 = \mu_n^{(k)}$
$\forall a > 0$.
\end{remark}

\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) and \(A\) be
organized by community such that the elements of blocks \(P^{(kl)}\) and
\(A^{(kl)}\) correspond to the edges between communities \(k\) and
\(l\).

\emph{Case \(k = l\)}. \(P^{(kk)}\) and \(A^{(kk)}\) represent
within-community edge probabilities and edges for community \(k\).\\
By definition, \(P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top\). This
implies that the singular value decomposition
\(P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top\) has one singular
value and one pair of singular vectors (\(P^{(kk)}\) is symmetric, so
the left and right singular vectors are identical). Then
\(\lambda^{(kk)} = \sigma_{kk} u^{(kk)}\).\\
Let \(\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top\) be the
singular value decomposition of \(A^{(kk)}\), and let
\(\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top\) be its
one-dimensional approximation. Define
\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding
approximation of \(\lambda^{(kk)}\).\\
Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the
adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
\(\lambda^{(kk)}\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).

>>>>

Define the subspace incoherence:
$$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$
Then $\forall k$, we have

@@

\begin{proof} From Theorem \ref{theorem3} we have that the subspaces
  $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
  are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
  \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
  \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
  $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
    $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
    0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
    \in V^{(-k)}$ and hence $\mu_n^{(k)} = 0$ as desired.
%   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
% must be orthogonal to each row of \(V^{(-k)}\) and hence
%\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
\end{proof}

@@

\begin{remark}
Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $\|v_n^{(i)}\| = 1$, 
which scales each $r_n^{(k)} \leq 1$. This is not strictly necessary for 
the proof of Theorem \ref{theorem5} since each $\mu_n^{(k)} = 0$, so as long 
as the $k^{th}$ community spans its subspace, $a r_n^{(k)} > 0 = \mu_n^{(k)}$ 
$\forall a > 0$.
\end{remark}

\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) be
organized by community such that \(P^{(k \ell)}\) denote the $n_k \times
n_{\ell}$ matrix obtained by keeping only the rows of $P$
corresponding to vertices in community $k$ and the columns of $P$
corresponding to vertices in community $\ell$. We define $A^{(k
  \ell)}$ analogously. Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
for the cases when $k = \ell$ versus when $k \not = \ell$.

\emph{Case \(k = l\)}. let $P^{(kk)} =
\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
decomposition of $P^{(kk)}$. We can then define
$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
singular value decomposition of \(A^{(kk)}\), and let
$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
best rank-one approximation of $A^{(kk)}$. Define
\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = \frac{(\log n_k)^c}{\sqrt{n_k}}.$$
Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.
